Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-12-11 manuscript/manuscript.tex
Improve density-based discretization algorithm

The numerical integration of an arbitrary continuous density function introduces
a new type of problem:
the integration error from using only a few nodes to discretize the density function.
The three dimensional adaptive discretization may help
to reduce this kind of error by adding more point masses in the radial direction.
However, it does not take into account the density function
and hence it is not well suited to fully perform this task.
Moreover, because the two dimensional adaptive discretization algorithm does not
subdivide the tesseroid on the radial dimension, it does not put any effort on lowering
this new kind of error.

We have developed a complementary discretization
algorithm that takes into account the variations of the density function.
This density-based discretization happens prior to the adaptive discretization described
in the previous section.
In short, the algorithm divides the tesseroid along the radial dimension at the
depths at which the \emph{maximum density variations} take place.

Consider an \emph{original} tesseroid with density given by the function $\rho(r')$.
Before the density-based discretization starts,
we normalise the density function to the range $[0, 1]$ as follows

@@

Once all smaller tesseroids satisfy Eq. \ref{eq:delta-density},
each one is subjected to the adaptive
discretization algorithm described earlier to calculate their gravitational effects.

On the first iteration, the ratio $L_r/L_r^\text{orig} = 1$ because the tesseroid being
divided is the original one.
For further iterations, the ratio will be progressively smaller than one as the
tesseroids get smaller.
This is intended to limit the number of divisions to the ones that will
significantly reduce the numerical error:
dividing a large tesseroid with a small $\text{max}\{ \Delta \rho(r') \}$ would
improve the integration accuracy more than dividing a small tesseroid with a
higher $\text{max}\{ \Delta \rho(r') \}$.
We also do not divide tesseroids with $L_r < 1$ mm because any integration errors would
be negligible due to the small mass of the tesseroid.

@@

\textit{Step 3:}
Apply a second-order GLQ to numerically compute the gravitational fields
(Eq.\ref{eq:glq-var-dens}) generated by each tesseroid of the set obtained in the
previous step.
Then sum these results in order to obtain the gravitational field of the original
tesseroid.

>>>>

The numerical integration of an arbitrary continuous density function introduces a new
type of problem: the integration error from using only a few nodes to account for the
variation of the density function.
The three dimensional adaptive discretization may help
to reduce this kind of error by adding more point masses in the radial direction.
However, it does not take into account the density function when subdividing the
tesseroid and hence it is not well suited to fully perform this task.
The two dimensional adaptive discretization algorithm does not subdivide the tesseroid
on the radial dimension, so it does not put any effort on lowering this new kind of
error.

We have developed a complementary discretization
algorithm that takes into account the variations of the density function.
This density-based discretization is applied prior to the adaptive discretization
described in the previous section.
In short, the algorithm divides the tesseroid along the radial dimension at the
depths at which the \emph{maximum density variations} take place.

Consider an \emph{original} tesseroid with a continuous density given by the function
$\rho(r')$.
Before the density-based discretization starts,
we normalise the density function to the range $[0, 1]$ as follows

@@

Once all smaller tesseroids satisfy Eq. \ref{eq:delta-density},
each one is subjected to the two dimensional adaptive
discretization algorithm described earlier to calculate their gravitational effects.

On the first iteration, the ratio $L_r/L_r^\text{orig} = 1$ because the tesseroid being
divided is the original one.
For further iterations, the ratio will be progressively smaller than one as the
tesseroids get smaller.
This is intended to limit the number of divisions to the ones that will
significantly reduce the numerical error:
dividing a large tesseroid with a small $\text{max}\{ \Delta \rho(r') \}$ would
improve the integration accuracy more than dividing a small tesseroid with a
higher $\text{max}\{ \Delta \rho(r') \}$.

@@

\textit{Step 3:}
Apply a second-order GLQ to numerically compute the gravitational fields
(Eq.\ref{eq:glq-var-dens}) generated by each tesseroid obtained in the
previous step.
Then sum these results in order to obtain the gravitational field of the original
tesseroid.
